---
title: "p8105_hw2_kd2942"
output: github_document
date: "2024-10-02"
---

#setting up markdown document
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

#loading libraries
```{r libraries}
library(tidyverse)
library(readxl)
```

**PROBLEM 1**
#read in Subway dataset
```{r}
subway_df=read.csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
subway_df=janitor::clean_names(subway_df)
```

#taking a look at the data 
```{r}
subway_df
skimr::skim(subway_df)
```

## 'select'
```{r}
subway_df = select(subway_df, line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada)
```

##converting entry from character to logical variable
```{r}
mutate(subway_df, entry = ifelse(subway_df$entry =="YES", TRUE, FALSE))
view(subway_df)
```

#writing about dataset
The data set contains the variables subway_df, line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, and ada. Data cleaning steps so far have included cleaning variable names, selecting certain variables, and turning entry into a logical variable. This data is not tidy, as there are missing data values and remaining inconsistencies between variables type (i.e. some route variables are integers while others are character variables).

##distinct stations
```{r}
subway_df_distinct = distinct(subway_df,station_name, line)
view(subway_df_distinct)
```
There are 465 distinct stations in the NYC subway. 

##ADA compliance
```{r}
filter(subway_df, ada =="TRUE")
subway_df
```
There are 468 stations that are ADA compliant.

##station exits and entrances without vending that allow entrance
```{r}
view(filter(subway_df, vending=="NO"))
view(filter(subway_df, vending == "NO", entry == "YES"))
```
There are 183 entrances/exits that do not have vending. Of the entrances and exits that do not have vending, 69 allow entry. 69/183 = 0.37705 or **37.705%**

##distinct trains serving A station and their ADA compliance
```{r eval=FALSE}
subway_df %>%
  col_types= cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")%>%
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route") %>% 
  filter(route == "A") %>%
  select(station_name, line) %>% 
  distinct()

subway_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>%
  distinct()
```
#Coded eval=false because I did not want knitting to be affected by my inability to figure out the above code.


**PROBLEM 2**
#read in Mr.Trash dataset
```{r}
trash_df = read_excel("./202409 Trash Wheel Collection Data.xlsx", sheet="Mr. Trash Wheel", na = c("NA", ".", ""))
trash_df
```

#clean Mr.trash dataset
```{r}
trash_df=janitor::clean_names(trash_df)
drop_na(trash_df,dumpster)
skimr::skim(trash_df)
```

#professor trash wheel dataset: read and clean
```{r}
professor_df = read_excel("./202409 Trash Wheel Collection Data.xlsx", sheet="Professor Trash Wheel", na = c("NA", ".", "")) 
professor_df=janitor::clean_names(professor_df)
drop_na(professor_df,dumpster)
```

#gwynnda trash wheel dataset: read and clean
```{r}
Gwynnda_df = read_excel("./202409 Trash Wheel Collection Data.xlsx", sheet="Gwynnda Trash Wheel", na = c("NA", ".", "")) 
Gwynnda_df=janitor::clean_names(Gwynnda_df)
drop_na(Gwynnda_df,dumpster)
```

#combining data sets
```{r}
mrtrashwheel = 
  readxl::read_excel("./202409 Trash Wheel Collection Data.xlsx", sheet="Mr. Trash Wheel", 
  mutate(wheel = "mrtrashwheel")

professortrashwheel = 
  readxl::read_excel("./202409 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel") |>
  mutate(movie = "professortrashwheel")

gwynnda = 
  readxl::read_excel("./202409 Trash Wheel Collection Data.xlsx", sheet="Gwynnda Trash Wheel") |>
  mutate(movie = "gwynnda")

trash_tidy = 
  bind_rows(mrtrashwheel, professortrashwheel, gwynnda) |>
  janitor::clean_names() 
```

